Oregon distributed lag model best fit degree of freedom

### 2017-11-06 r_scripts/analysis/model_dl_spline_knots.Rmd

---
title: "Oregon_manuscript"
author: "Jingyang Liu"
date: "2017/10/30"
output: html_document
---

@@

# server
library(dplyr)
library(data.table)
library(readxl) 
library(readr)

@@

## Overview

**1. Introduction: **
This data set of sample population for our research come from the Oregon All Payer All Claims Database (APAC) in the year 2013, which recorded the health care data for Oregon’s insured populations. This database included our main researched aspects, which are individuals with both disease and pharmacy claims. The disease codes were International Classification of Diseases, Clinical Modification (ICD-9-CM) diagnoses codes, and the pharmacy codes were National Drug Codes (NDC).

We already found an association with wildfire smoke and respiratory outcomes in Washington state in 2012 wildfire season using a novel estimate of smoke concentration, geographically weighted ridge regression (GWR). We wanted to know if smoke impacts other health outcomes, specifically respiratory rescue medications (beta 2 agonists) (abbreviate to SABA). However, we did not have the specific NDC for asthma or respiratory, then we transfered to find the SABA instead and made categories for different diseases.

**2. Research question: **
We evaluated the association between smoke concentrations using the GWR method and SABA refills in Oregon state during the 2013 wildfire season.

We had previously shown GWR to be a good estimate of wildfire smoke concentrations, and showed associations with respiratory outcomes (including asthma) in Washington state. We want to replicate results for Oregon and also look at beta-2 agonist pharmacy fills.

**3.Methods: **

We limited the population to patients whose admission date are during the wildfire season (May 1 to September 30, 2013) within Oregon State, and checked the zip codes of these patients. We further limited the population who were billed for filling a SABA prescription, which we used as a proxy for use. Finally, we limited to the subjects of the unique patient’s first visit to reduce the potential bias of counting a event multiple times.

In these comparisons, we examined GWR method of smoke PM2.5 estimations and associations with medication outcomes using a time-stratified case-crossover study design. Pharmacy outcomes for a patient with a primary diagnosis of all pharmacy records from Oregon state were recorded, and we also had APAC data for people who were billed for a cardiopulmonary event that we could also link the disease to the pharmacy claims, in which their dates of admission (index time) were also identified. We then created counterfactual observations for each patient on the same day of the week during the entire wildfire season for Oregon.

1. Data cleaning for health outcomes

There were about 77 million claims in the total data set, of which 23 million were pharmacy records denoted by the NDC variable. So we chose the claims in Oregon State with ndc existing, which was about 23 million claims. Also, the POS (Place of service code) all showed "pharmacy" type, comparing to the previous health disease outcomes we limited the "emergency visit" type and the "urgent care" type.

For ndc, we chose the people who were billed for filling a beta2 agonist prescription, which we used as a proxy for use to form the sample. Because the ndc csv file had more information than what we need, we filtered the ndc file by SQLite Studio to a file only containing the beta 2 agonists medication information for filtering patients.

After choosing the claims with beta 2 agonists, the data set was reduced to about 232 thousand observations. After first recorded pharmacy fill and limit the wildfire season dates (May 1st, 2013 to September 30th, 2013), the data set was the final data set for analysis, which contained 24,538 beta-2 agonist records.

```{r import filtered original data and make index, eval=FALSE, echo=FALSE}
### Import whole data set ------------------------------------------------------
read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
start_time <- Sys.time()
oregon_df <- fread(read_path, sep = "|", showProgress = T) 
stop_time <-  Sys.time() - start_time 
# time it took
stop_time # 7 mins

### Import NDC data set --------------------------------------------------------
read_path <- paste0("../../../data/data_new/oregon_ndc.csv")
oregon_ndc <- read_csv(read_path)
dim(oregon_ndc) # 23330714 obs

## Import Asthma NDC table
read_path2 <- paste0("../../../data/data_new/ndc_2013_beta_2_agonists.csv")
inhaler_ndc <- read_csv(read_path2, col_names = FALSE)

colnames(inhaler_ndc) <- c("ndc_code", "brand_name", "generic_product_name", 
                          "route", "category", "drug_id")

# convert to vector
ndc_codes <- as.vector(as.matrix(inhaler_ndc$ndc_code))

oregon_beta <- oregon_ndc %>%
  filter(ndc %in% ndc_codes) # 232665, unique personkey 83787

oregon_beta_df <- oregon_beta %>% 
  group_by(personkey) %>% 
  arrange(personkey, fromdate, line) %>% 
  mutate(num_visit = dense_rank(fromdate)) %>%
  arrange(personkey, num_visit) %>% 
  select(personkey, clmid, num_visit) %>% 
  unique() %>% 
  full_join(oregon_beta, by = c("personkey", "clmid")) %>% 
  arrange(personkey, clmid, line, fromdate) %>% 
  filter(row_number() == 1) %>%
  # add new transverted from date
  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) # 83787 (same)

oregon_df_ndc <- oregon_df %>%
  filter(personkey %in% oregon_beta$personkey) # 8909892

### Import ICD9 code
icd9_key <- read_excel("../../../data/data_original/CMS32_DESC_LONG_SHORT_DX.xlsx") %>% 
  # rename the terrible variable names
  select(dx_code = 1, long_desc = 2, short_desc = 3)

# sort by icd9 code add row variable 
icd9_key$X <- NULL
icd9_key <- arrange(icd9_key, dx_code) %>%
  mutate(n = as.numeric(row.names(icd9_key)))

### without ICD-9
or_ndc_no_icd <- oregon_df_ndc %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="" )

length(or_ndc_no_icd$personkey) # 3656720 claims
length(unique(or_ndc_no_icd$personkey)) # 83787

or_no_ndc_no_icd <- oregon_df %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 1057022 missing in total data set

or_ndc_all_miss <- oregon_df_ndc %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 20269 missing in total data set

or_ndc_all_miss_id <- or_ndc_asthma_df %>%
  filter(personkey %in% unique(or_ndc_copd_df$personkey)) %>% # 7279 person
  select(personkey)

oregon_ndc_dx <- oregon_ndc %>%
    filter(dx1!="*NULL*" |
         dx2!="*NULL*" |
         dx3!="*NULL*" |
         dx4!="*NULL*" |
         dx5!="*NULL*" |
         dx6!="*NULL*" |
         dx7!="*NULL*" |
         dx8!="*NULL*" |
         dx9!="*NULL*" |
         dx10!="*NULL*" |
         dx11!="" |
         dx12!="" |
         dx13!="" ) # 0 claims

### asthma
which(icd9_key$dx_code == '49300') # start of asthma is row 5206
which(icd9_key$dx_code == '49392') # end of asthma is row 5219

# limit just to asthma code and just the diagnosis column
icd9_check <- icd9_key %>%
  filter(n >= 5206 & n <= 5219) 
icd9_check

asthma_icd9 <- filter(icd9_key, n >= 5206 & n <= 5219) %>%
  select(dx_code)
# convert to vector
asthma_icd9 <- as.vector(as.matrix(asthma_icd9))

# now can I make a new variable, asthma1, that indicates an asthma claim?
or_ndc_asthma_df <- oregon_df_ndc %>%
  filter(dx1 %in% asthma_icd9 |
         dx2 %in% asthma_icd9 |
         dx3 %in% asthma_icd9 |
         dx4 %in% asthma_icd9 |
         dx5 %in% asthma_icd9 |
         dx6 %in% asthma_icd9 |
         dx7 %in% asthma_icd9 |
         dx8 %in% asthma_icd9 |
         dx9 %in% asthma_icd9 |
         dx10 %in% asthma_icd9 |
         dx11 %in% asthma_icd9 |
         dx12 %in% asthma_icd9 |
         dx13 %in% asthma_icd9) # 375451, unique personkey: 45165

### COPD
copd_icd9 <- c('490', '4910','4911','49120','49121','49122','4918','4919', '4920',
               '4928', '4940', '4941', '496')

# now can I make a new variable, asthma1, that indicates an asthma claim?
or_ndc_copd_df <- oregon_df_ndc %>%
  filter(dx1 %in% copd_icd9 |
         dx2 %in% copd_icd9 |
         dx3 %in% copd_icd9 |
         dx4 %in% copd_icd9 |
         dx5 %in% copd_icd9 |
         dx6 %in% copd_icd9 |
         dx7 %in% copd_icd9 |
         dx8 %in% copd_icd9 |
         dx9 %in% copd_icd9 |
         dx10 %in% copd_icd9 |
         dx11 %in% copd_icd9 |
         dx12 %in% copd_icd9 |
         dx13 %in% copd_icd9) # 361738, unique personkey: 18626

### asthma or copd
or_or_asthma_copd <- oregon_df_ndc %>%
  filter(dx1 %in% asthma_icd9 |
         dx2 %in% asthma_icd9 |
         dx3 %in% asthma_icd9 |
         dx4 %in% asthma_icd9 |
         dx5 %in% asthma_icd9 |
         dx6 %in% asthma_icd9 |
         dx7 %in% asthma_icd9 |
         dx8 %in% asthma_icd9 |
         dx9 %in% asthma_icd9 |
         dx10 %in% asthma_icd9 |
         dx11 %in% asthma_icd9 |
         dx12 %in% asthma_icd9 |
         dx13 %in% asthma_icd9 |
         dx1 %in% copd_icd9 |
         dx2 %in% copd_icd9 |
         dx3 %in% copd_icd9 |
         dx4 %in% copd_icd9 |
         dx5 %in% copd_icd9 |
         dx6 %in% copd_icd9 |
         dx7 %in% copd_icd9 |
         dx8 %in% copd_icd9 |
         dx9 %in% copd_icd9 |
         dx10 %in% copd_icd9 |
         dx11 %in% copd_icd9 |
         dx12 %in% copd_icd9 | 
         dx13 %in% copd_icd9) # 722107 , unique personkey: 56512

# 45165+ 18626 - 56512
# [1] 7279 on both asthma and copd

### Both asthma and copd
length(unique(or_or_asthma_copd$personkey))
length(unique(or_ndc_asthma_df$personkey))
length(unique(or_ndc_copd_df$personkey))

or_with_icd <- oregon_df_ndc %>%
  filter(!dx1=="*NULL*" |
         !dx2=="*NULL*" |
         !dx3=="*NULL*" |
         !dx4=="*NULL*" |
         !dx5=="*NULL*" |
         !dx6=="*NULL*" |
         !dx7=="*NULL*" |
         !dx8=="*NULL*" |
         !dx9=="*NULL*" |
         !dx10=="*NULL*" |
         !dx11=="" |
         !dx12=="" |
         !dx13=="" )

asthma_copd_both_id <- or_ndc_asthma_df %>%
  filter(personkey %in% unique(or_ndc_copd_df$personkey)) %>% # 7279 person
  select(personkey) %>%
  unique() %>%
  mutate(index="both")

### Only asthma
asthma_only_id <- or_ndc_asthma_df %>%
  filter(!personkey %in% asthma_copd_both_id$personkey) %>% # 37886 person
  select(personkey) %>%
  unique() %>%
  mutate(index="asthma")

### Only copd
copd_only_id <- or_ndc_copd_df %>%
  filter(!personkey %in% asthma_copd_both_id$personkey) %>% # 11347 person
  select(personkey) %>%
  unique() %>%
  mutate(index="copd")

### Other
other_disease_id <- oregon_df_ndc %>%
  filter(!personkey %in% or_or_asthma_copd$personkey) %>% # 27275 person
  select(personkey) %>%
  unique() %>%
  mutate(index="other")

id_index <- bind_rows(asthma_copd_both_id, asthma_only_id, copd_only_id, other_disease_id)

## Join the index with beta 2 ndc claims
oregon_beta_index <- oregon_beta %>%
  left_join(id_index, by="personkey")
summary(as.factor(oregon_beta_index$index))

write_path3 <- paste0('../../../data/data_new/',
                     'oregon_beta_ndc_index.csv')
write_csv(oregon_beta_index, write_path3)

## Join with original data and indicator for 4 categories
oregon_ndc_index <- oregon_df_ndc %>%
  left_join(id_index, by="personkey")
summary(as.factor(oregon_ndc_index$index))

write_path4 <- paste0('../../../data/data_new/',
                     'oregon_beta_ndc_index_full.csv')
write_csv(oregon_ndc_index, write_path4)

### Check
or_ndc_asthma_copd_sum <- oregon_ndc_index %>%
  mutate(asthma1 = ifelse(dx1 %in% asthma_icd9, 1, 0),
         asthma2 = ifelse(dx2 %in% asthma_icd9, 1, 0),
         asthma3 = ifelse(dx3 %in% asthma_icd9, 1, 0),
         asthma4 = ifelse(dx4 %in% asthma_icd9, 1, 0),
         asthma5 = ifelse(dx5 %in% asthma_icd9, 1, 0),
         asthma6 = ifelse(dx6 %in% asthma_icd9, 1, 0),
         asthma7 = ifelse(dx7 %in% asthma_icd9, 1, 0),
         asthma8 = ifelse(dx8 %in% asthma_icd9, 1, 0),
         asthma9 = ifelse(dx9 %in% asthma_icd9, 1, 0),
         asthma10 = ifelse(dx10 %in% asthma_icd9, 1, 0),
         asthma11 = ifelse(dx11 %in% asthma_icd9, 1, 0),
         asthma12 = ifelse(dx12 %in% asthma_icd9, 1, 0),
         asthma13 = ifelse(dx13 %in% asthma_icd9, 1, 0),
         # sum up the asthma indicators
         asthma_sum = (asthma1 + asthma2 + asthma3 + asthma4 + asthma5 + asthma6 + asthma7 + asthma8 + asthma9 + asthma10 + asthma11 + asthma12 + asthma13),
         asthma_dx = ifelse(asthma_sum>0, asthma_sum, 0)
         ) %>% # end of mutate asthma
  mutate(copd1 = ifelse(dx1 %in% copd_icd9, 1, 0),
         copd2 = ifelse(dx2 %in% copd_icd9, 1, 0),
         copd3 = ifelse(dx3 %in% copd_icd9, 1, 0),
         copd4 = ifelse(dx4 %in% copd_icd9, 1, 0),
         copd5 = ifelse(dx5 %in% copd_icd9, 1, 0),
         copd6 = ifelse(dx6 %in% copd_icd9, 1, 0),
         copd7 = ifelse(dx7 %in% copd_icd9, 1, 0),
         copd8 = ifelse(dx8 %in% copd_icd9, 1, 0),
         copd9 = ifelse(dx9 %in% copd_icd9, 1, 0),
         copd10 = ifelse(dx10 %in% copd_icd9, 1, 0),
         copd11 = ifelse(dx11 %in% copd_icd9, 1, 0),
         copd12 = ifelse(dx12 %in% copd_icd9, 1, 0),
         copd13 = ifelse(dx13 %in% copd_icd9, 1, 0),
         # sum up the pneumonia indicators
         copd_sum = (copd1 + copd2 + copd3 + copd4 +copd5 + copd6 + copd7 + copd8 + copd9 +copd10 + copd11 + copd12 + copd13),
         copd_dx = ifelse(copd_sum>0, copd_sum, 0)
  ) #end of mutate of copd

or_ndc_check <- or_ndc_asthma_copd_sum %>%
  select(personkey, index, asthma_dx, copd_dx)
summary(as.factor(or_ndc_check$asthma_dx))

or_asthma <- or_ndc_asthma_copd_sum %>%
  filter(index=="asthma")
or_copd <- or_ndc_asthma_copd_sum %>%
  filter(index=="copd")
or_both <- or_ndc_asthma_copd_sum %>%
  filter(index=="both")
or_other <- or_ndc_asthma_copd_sum %>%
  filter(index=="other")

summary(as.factor(or_other$dx1))

## check for one sample
summary(as.factor(or_asthma$personkey))
summary(as.factor(or_copd$personkey))
summary(as.factor(or_both$personkey))
summary(as.factor(or_other$personkey))

or_asthma_sample <- or_asthma %>%
  filter(personkey == 10101244) 
which(or_asthma_sample$asthma_dx!=0)

or_copd_sample  <- or_copd %>%
  filter(personkey == 13564804)

or_both_sample  <- or_both %>%
  filter(personkey == 12979399)

or_other_sample  <- or_other %>%
  filter(personkey == 10289787)

or_asthma_miss <- or_asthma %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 89482 claims missing in asthma data set

or_copd_miss <- or_copd %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 18941 claims missing in copd data set

or_both_miss <- or_both %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 17963 claims missing in both data set

or_other_miss <- or_other %>%
  filter(dx1=="*NULL*" &
         dx2=="*NULL*" &
         dx3=="*NULL*" &
         dx4=="*NULL*" &
         dx5=="*NULL*" &
         dx6=="*NULL*" &
         dx7=="*NULL*" &
         dx8=="*NULL*" &
         dx9=="*NULL*" &
         dx10=="*NULL*" &
         dx11=="" &
         dx12=="" &
         dx13=="") %>%
  filter(ndc=="" |
         ndc=="*NULL*") # 46971 claims missing in other data set (5848 personkey)

or_other_not_miss <- or_other %>%
  filter(dx1!="*NULL*" |
         dx2!="*NULL*" |
         dx3!="*NULL*" |
         dx4!="*NULL*" |
         dx5!="*NULL*" |
         dx6!="*NULL*" |
         dx7!="*NULL*" |
         dx8!="*NULL*" |
         dx9!="*NULL*" |
         dx10!="*NULL*" |
         dx11!="" |
         dx12!="" |
         dx13!="" |
         ndc!="" |
         ndc!="*NULL*") # 20269 missing in total data set

length(unique(or_other_not_miss$personkey)) # [1] 27275

dim(or_asthma_miss)
dim(or_copd_miss)
dim(or_both_miss)
dim(or_other_miss)

sum(or_ndc_all_miss_id %in% unique(or_other$personkey))

# write_path <- paste0('../../../data/data_new/',
#                      'oregon_asthma_sample.csv')
# write_csv(or_asthma_sample, write_path)

# write_path2 <- paste0('../../../data/data_new/',
#                      'oregon_ndc_with_index.csv')

# write_csv(or_ndc_asthma_copd_sum, write_path2)

```{r choose ndc, eval=FALSE, echo=FALSE}

# limit to the unique personkey different days' first visit
oregon_asthma_df <- oregon_asthma %>% 
  group_by(personkey) %>% 
  arrange(personkey, fromdate, line) %>% 
  mutate(num_visit = dense_rank(fromdate)) %>%
  arrange(personkey, num_visit) %>% 
  select(personkey, clmid, num_visit) %>% 
  unique() %>% 
  full_join(oregon_asthma, by = c("personkey", "clmid")) %>% 
  arrange(personkey, clmid, line, fromdate) %>% 
  filter(row_number() == 1) %>%
  # add new transverted from date
  mutate(dates = as.Date(fromdate, "%m/%d/%Y")) %>%
  filter(dates >= '2013-05-01' & 
         dates <= '2013-09-30') # 24538

write_path2 <- paste0("../../../data/data_new/oregon_asthma_ndc.csv")
write_csv(oregon_asthma_df, write_path2)

2. Exposure methods for smoke model

For our research question, we focused mainly on the GWR method (geographically weighted ridge regression). For Geo-Weighted Regression model, with 'smk' designator, we subtracted off the 'Background' estimates of smoke, which were the monthly averages of PM2.5 for a given grid. However, the 'Background' given was the mean of total period from May 1st to Sep 30th 2013 so we had to assign the mean smoke estimates for every day. For patients who had a index time before May 1, 2013 or after September 30, 2013, their referent observations before or after these dates would be excluded because we were not able to assign estimates of PM2.5 to these observations by atmospheric department.

3. Analytic method

To analyze the result, we used the conditional logistic regression model using the survival package in R. Each conditional logistic regression model accounted for the subject, and adjusted for temperature from the GWR smoke model. The conditional logistic regression model  represented the change in risk of an pharmacy event associated with a short-term unit increase in exposure, and could be calculated as an average difference between exposure at the index time and a weighted average of exposure at all times in the referent window. (???)

We also used distributed lag model because we still wondered the association of the pharmacy fills with some lag days. For medication analysis, people could get the pharmacy fills one or some days later than the day they find they have the disease. Then we made 0 (the original day) to 7 lags (one week) to see association of pharmacy fills with different lagged days.

```{r import original data, eval=FALSE, echo=FALSE}
### Import whole data set ------------------------------------------------------
read_path <- paste0("../../../data/data_original/gan_episodes_of_care.txt")
start_time <- Sys.time()
oregon_df <- fread(read_path, sep = "|", showProgress = T) 
stop_time <-  Sys.time() - start_time 
# time it took
stop_time # 7 mins

## Basic cleaning
start_time <- Sys.time()
oregon_df2 <- oregon_df %>% 
  filter((!ndc=="")&(!ndc=="*NULL*")) %>% 
  # filter State is Oregon
  filter(STATE=="OR")
stop_time <-  Sys.time() - start_time 
# time it took
stop_time # 40 secs

write_path <- paste0('../../../data/data_new/',
                     'oregon_ndc.csv')
write_csv(oregon_df2, write_path)
```

Time-Stratified Case-Crossover (Population-Weighted PM~2.5~ Assigned by Zip Code)
*referents same day of week within fire season May - September*

The time-stratified case-crossover design used subjects as their own controls. For each pharmacy claim, we also created counterfactual observations on the the same day of the week over the fire season for each subject to indicate the outcome for this day was "1". Then we made the dates week by week before this date, and after this date in the whole wildfire season, and indicated these dates' outcomes were "0". Then we could use these outcomes for the following regression model building.

```{r casecrossover, eval=FALSE, echo=FALSE}
### Casecrossover study
read_path3 <- paste0("../../../data/data_new/medication/oregon_asthma_ndc.csv")
oregon_asthma_df <- read_csv(read_path3)

outcome_id <- oregon_asthma_df %>%
  # arrange with dates
  arrange(dates) %>%
  mutate(id = seq(1, nrow(.), by = 1))

# create dataset to populate
id_date_df <- data_frame()

# begin second loop to create counterfactual observations for each case subject
for (k in 1:nrow(outcome_id)){

  # find the replicate times of weeks
  dates_l <- outcome_id[[k,74]] 
  n1 <- 0
  d=as.Date("2013-05-01")
  i=1
  while (dates_l >= "2013-05-01"){
    dates_l <- dates_l - 7
    d[i] = dates_l
    i = i+1
    n1 = n1+1
  }
  d[1:n1-1] # shows character(0) when the first week
  n1-1

  dates_l <- outcome_id[[k,74]] 
  n2=0
  e=as.Date("2013-09-30")
  j=1
  while (dates_l <= "2013-09-30"){
    dates_l <- dates_l + 7
    e[j]=dates_l
    j=j+1
    n2 = n2 + 1
  }
  e[1:n2-1] # shows character(0) when the last week
  n2-1

  # replicate covariates length of counterfactual dates
  # and make conuterfactual dates
  if (n1==1){
    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
    cov_df$dates <- c(outcome_id[[k,74]], e[1:n2-1])
  } else if (n2==1){
    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
    cov_df$dates <- c(outcome_id[[k,74]], d[1:(n1-1)])
  }else{
    cov_df <- do.call("bind_rows", replicate(n1+n2-1, outcome_id[k,],simplify = F))
    cov_df$dates <- c(outcome_id[[k,74]], d[1:(n1-1)], e[1:n2-1])
  }

  # bind unique id and date of the year with covariates
  id_date <- bind_cols(cov_df)
  # iteration which binds rows of unique ids
  id_date_df <- bind_rows(id_date_df, id_date)
}

outcome_casecross <- id_date_df %>%
  mutate(outcome = ifelse(dates == as.Date(fromdate, "%m/%d/%Y"), 1, 0)) %>%
  arrange(id, dates) # order by id and date

outcome_casecross <- id_date_df %>%
  mutate(outcome = ifelse(dates == as.Date(fromdate, "%m/%d/%Y"), 1, 0)) %>%
  arrange(personkey, dates) # order by id and date

write_path2 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv")
write_csv(outcome_casecross, write_path2)

Also, to get the time stratified data set, we created the lag variables that take smoke values from the previous days for zip codes, joined with the zip code level population-weighted pm to the data set, and create age, sex, month and season index.

```{r time stratified zip for casecrossover, eval=FALSE, echo=FALSE}
### Time stratified ------------------------------------------------------------
# read in zipcode level populatoin-weighted pm
read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')

zip_smoke <- read_csv(read_path5) # 63801 rows

# descriptives of the two smoke datasets
summary(zip_smoke)

# Zipcode PM2.5 estimates
# create lag variables that take smoke values from n previous days for zipcodes
zip_smoke_w_lag <- zip_smoke %>% arrange(ZIPCODE, date) %>%
  # group by zipcode
  group_by(ZIPCODE) %>% 
  # wrf
  mutate(wrf_f_pm_lag1 = lag(wrf_f_pm, 1, order_by = ZIPCODE), 
         wrf_f_pm_lag2 = lag(wrf_f_pm, 2, order_by = ZIPCODE),
         wrf_f_pm_lag3 = lag(wrf_f_pm, 3, order_by = ZIPCODE),
         wrf_f_pm_lag4 = lag(wrf_f_pm, 4, order_by = ZIPCODE),
         wrf_f_pm_lag5 = lag(wrf_f_pm, 5, order_by = ZIPCODE),
         wrf_f_pm_lag6 = lag(wrf_f_pm, 6, order_by = ZIPCODE),
         wrf_f_pm_lag7 = lag(wrf_f_pm, 7, order_by = ZIPCODE),
         # wrf no fire lag
         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = ZIPCODE),
         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = ZIPCODE),
         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = ZIPCODE),
         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = ZIPCODE),
         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = ZIPCODE),
         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = ZIPCODE),
         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = ZIPCODE),
         # wrf_smk_pm
         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = ZIPCODE),
         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = ZIPCODE),
         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = ZIPCODE),
         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = ZIPCODE),
         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = ZIPCODE),
         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = ZIPCODE),
         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = ZIPCODE),
         # geo weighted pm
         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = ZIPCODE),
         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = ZIPCODE),
         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = ZIPCODE),
         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = ZIPCODE),
         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = ZIPCODE),
         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = ZIPCODE),
         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = ZIPCODE),
         # krig pm
         krig_pm_lag1 = lag(krig_pm, 1, order_by = ZIPCODE),
         krig_pm_lag2 = lag(krig_pm, 2, order_by = ZIPCODE),
         krig_pm_lag3 = lag(krig_pm, 3, order_by = ZIPCODE),
         krig_pm_lag4 = lag(krig_pm, 4, order_by = ZIPCODE),
         krig_pm_lag5 = lag(krig_pm, 5, order_by = ZIPCODE), 
         krig_pm_lag6 = lag(krig_pm, 6, order_by = ZIPCODE),
         krig_pm_lag7 = lag(krig_pm, 7, order_by = ZIPCODE), 
         # background pm
         background_pm_lag1 = lag(background_pm, 1, order_by = ZIPCODE),
         background_pm_lag2 = lag(background_pm, 2, order_by = ZIPCODE),
         background_pm_lag3 = lag(background_pm, 3, order_by = ZIPCODE),
         background_pm_lag4 = lag(background_pm, 4, order_by = ZIPCODE),
         background_pm_lag5 = lag(background_pm, 5, order_by = ZIPCODE), 
         background_pm_lag6 = lag(background_pm, 6, order_by = ZIPCODE),
         background_pm_lag7 = lag(background_pm, 7, order_by = ZIPCODE), 
         # geo_smk_pm 
         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = ZIPCODE),
         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = ZIPCODE),
         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = ZIPCODE),
         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = ZIPCODE),
         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = ZIPCODE),
         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = ZIPCODE),
         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = ZIPCODE),
         # krig smk pm
         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = ZIPCODE),
         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = ZIPCODE),
         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = ZIPCODE),
         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = ZIPCODE),
         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = ZIPCODE),
         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = ZIPCODE),
         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = ZIPCODE),
         # temp
         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = ZIPCODE),
         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = ZIPCODE),
         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = ZIPCODE),
         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = ZIPCODE),
         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = ZIPCODE), 
         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = ZIPCODE),
         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = ZIPCODE)) %>% 
  # ungroup by zip
  ungroup(ZIPCODE) %>% 
  # attach a zip indicator for each smoke variable
  setNames(paste(colnames(.), "zip", sep="_"))  %>% 
  # remove the '_zip' from the zipcode and date variable 
  rename(ZIPCODE = ZIPCODE_zip, date = date_zip)

read_path4 <- paste0("../data_new/medication/oregon_ndc_casecrossover.csv")
# read_path4 <- paste0("../../../data/data_new/oregon_ndc_casecrossover.csv") # for server
or_disease <- read_csv(read_path4)

### try join
colnames(or_disease)[24] <- c("ZIPCODE")
colnames(or_disease)[74] <- c("date")

>>>>

---
title: "Oregon distributed lag model best fit degree of freedom"
author: "Jingyang Liu"
date: "2017/11/6"
output: html_document
---

@@

@@

Load data
